---
title: "data-reanalysis-assignment"
author: "Brandon Güell"
date: "10/20/2017"
output: html_document
---
#Data Reanalysis Assignment

The original study (Gomez-Mestre I, Warkentin KM. 2007. To hatch and hatch not: Similar selective trade-offs but different responses to egg predators in two closely related, syntopic treefrogs. Oecologia 153:197–206.) examined differences in selective trade-offs and responses to predation in two closely related and syntopic treefrog species: Agalychnis callidryas and A. spurrelli. The study performed field surveys to collect data on natural spontaneous hatching timings and natural predation on egg clutches of the two species. Authors also performed predation experiments using snakes for predation on eggs and fish for predation on hatchlings.

Their specific null hypotheses were outlined explicitly, "(1) both species experience similar levels of egg predation; (2) they have the same pattern of undisturbed or spontaneous hatching; (3) their escape hatching responses to snake attack and flooding are similar; and (4) early hatching increases vulnerability to aquatic predators in A. spurrelli, as it does in A. callidryas."

This paper used multiple data sets from various authors. All data was collected in the field, but in different years and by different authors. They performed a suite of statistical analysis, but the main ones were as follows:

A) "The overall proportion of monitored clutches predated in the field, escape hatching success in snake predation experiments, and proportion of hatchlings surviving Wsh predation were all analysed using generalized linear models with an underlying binomial distribution and a logit link function in PROC GENMOD."

B) "We performed a log-linear analysis of the number of clutches of each species that were predated and undisturbed on each of 16 nights when both were present. Model parameters were estimated through maximum likelihood using PROC CATMOD, assuming an underlying Poisson distribution."

C) "We tested for effects of embryo development on escape success in snake attacks using generalized linear models."

D) "The hatching response of A. spurrelli to underwater submergence was tested against its unflooded controls by Wtting a Cox regression model using PROC PHREG"

E) "The effects of hatching age on the proportion of hatchlings surviving in the fish predation experiments was analysed with a generalized linear model, including fish size as a covariate."

I will be reanalyzing data for which I gathered the original data set for. I will reproduce analysis for:

1) A: GLMs on field predation observations, snake predation experiments, and fish predation experiments; including the corresponding data manipulation, data summarization and descriptive statistical analysis, and visualization of each.

2) C: GLM to test for effects of age on escape success

3) E: GLM to test for effects of age and fish size on hatchling survival

So lets get started!

First we'll load several cvs files from the web for hte different chapters of this reanalysis project! Here's the first!
```{r}
library(curl) #load necessary package for uploading data from Github

f <- curl("https://raw.githubusercontent.com/bguell/data-reanalysis-assignment/master/undisturbed%20raw.csv") #read in csv file from Github
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d) #see data and column names
d

```
Now that the raw data for field surveys of undisturbed clutches are loaded, lets begin with some basic data summarization which are also found in the paper.

First, lets look at some basic summary statistics of clutch size for both species. We'll need to find the mean, standard deviations, and range of the clutch sizes for each:
```{r}
#lets start with callidryas:
range(d$number_of_eggs[d$species=="callidryas"]) #find range of number of eggs subsetting for only callidryas species
mean(d$number_of_eggs[d$species=="callidryas"]) #find the mean of number of eggs
sd(d$number_of_eggs[d$species=="callidryas"]) #find the standard deviation of number of eggs

#or using dplyr for 2 out of 3 summaries:
library(dplyr)
d %>% summarise(mean= mean(number_of_eggs), SD= sd(number_of_eggs)) # mean of prop_hatch by day!
#then the range separately
range(d$number_of_eggs[d$species=="callidryas"]) #find range of number of eggs subsetting for only callidryas species

#and the same for spurrelli:
range(d$number_of_eggs[d$species=="spurrelli"]) #find range of number of eggs for only callidryas species
mean(d$number_of_eggs[d$species=="spurrelli"]) #find the mean of number of eggs
sd(d$number_of_eggs[d$species=="spurrelli"]) #find the standard deviation of number of eggs

#or using dplyr for 2 out of 3 summaries:
d %>% summarise(mean= mean(number_of_eggs), SD= sd(number_of_eggs)) # mean of prop_hatch by day!
#then the range separately
range(d$number_of_eggs[d$species=="spurrelli"]) #find range of number of eggs subsetting for only callidryas species

```
Next, lets look at the overall proportion of predated clutches for each species and compare them using a test of equal proportions:
```{r}
#so, since the data sheet I was given doesn't have info about the total number of observed clutches, we need to calculate the total N for each species. We can back calculate this since we know the proportion of predated clutches for each species from the data reported in the paper, 0.40 nad 0.48 for callidryas and spurrelli respectively.
length(d$species[d$species=="callidryas"]) #number of predated clutches for callidryas
Ncalidryas= 182 *.40
Ncalidryas

length(d$species[d$species=="spurrelli"]) #number of predated clutches for spurrelli
Nspurrelli= 19 *.48
Nspurrelli

#then incorporating these numbers into a two sample proportion test:
prop.test(x=c(73,9), n=c(182,19)) #clearly the proportions predated for each species were not significaly different from each other
```

Next let's make some figures to visualize this data! We'll first need to seriously wrangle this data since there are multiple colums for observations which should be in just one... lets use dplyr and tidyr!!! yay data clean up time!
```{r}
library(dplyr)
d= select(d, 1:16) #only need to plot days 4-9
dd=gather(data = d, day, prop_hatch, D4.am:D9.pm, -year, -species, -clutch, -number_of_eggs) # accumulate all observations into one column named prop_hatch
dd

str(dd) #thats more like it! great!

#but lets separate the Day # from the am and pm if we can...
dd = separate(dd, day, c("day", "time_of_day"))
dd
str(dd) #wow fuck yes! that worked so freaking well! yay me!

```

```{r}
#now lets get species specific data so we can plot these as separate figures by species
ddc = filter(dd, species =="callidryas")
ddc
dds = filter(dd, species == "spurrelli")
dds
```

Now lets get some summary statistics to actually plot!
```{r}
ddc= ddc %>% group_by(day, time_of_day) #group by day colum to summarize by day!

mphc=ddc %>% summarise(mean_prop= mean(prop_hatch), SE= sd(prop_hatch)/sqrt(length(prop_hatch))) # mean of prop_hatch by day!
mphc

mphc= unite(mphc, "day_time", c("day", "time_of_day")) #and now combine the day and time of day colums back for easier plotting
mphc

mphc=mutate(mphc, time_of_day = c("am", "pm","am", "pm","am", "pm","am", "pm","am", "pm","am", "pm")) # add column to distingusih am and pm
mphc

dds= dds %>% group_by(day, time_of_day) #group by day colum to summarize by day!

mphs=dds %>% summarise(mean_prop= mean(prop_hatch), SE= sd(prop_hatch)/sqrt(length(prop_hatch))) # mean of prop_hatch by day!
mphs

mphs= unite(mphs, "day_time", c("day", "time_of_day")) #and now combine the day and time of day colums back for easier plotting
mphs

mphs=mutate(mphs, time_of_day = c("am", "pm","am", "pm","am", "pm","am", "pm","am", "pm","am", "pm")) # add column to distingusih am and pm
mphs


```
Excellent! Now lets go ahead and plot those to see what it looks like!
```{r}
library(ggplot2)

gc= ggplot(data=mphc, aes(x=day_time, y=mean_prop)) + geom_bar(stat="identity", aes(fill=time_of_day)) #use barplots w/time of day filled w/color
gc= gc + geom_errorbar(aes(ymin=mean_prop-SE, ymax=mean_prop+SE), width=.2, position=position_dodge(.9), color="black") #add black error bars
gc= gc + labs(x= "Age (days)", y= "Sponanteous hatching (proportion hatched eggs)") #make labels for axis
gc= gc + theme_bw(base_size = 11, base_family = "Times") #make theme in Times New Roman and b/w
gc= gc + annotate("text", x = 2, y = 0.5, label = "(b) A. callidryas", fontface= "italic") #add italicized text to figure!
gc

#and again same for spurrelli
gs= ggplot(data=mphs, aes(x=day_time, y=mean_prop)) + geom_bar(stat="identity", aes(fill=time_of_day))
gs= gs + geom_errorbar(aes(ymin=mean_prop-SE, ymax=mean_prop+SE), width=.2, position=position_dodge(.9), color="black") 
gs= gs + labs(x= "Age (days)", y= "Sponanteous hatching (proportion hatched eggs)") 
gs= gs + theme_bw(base_size = 11, base_family = "Times")
gs= gs + annotate("text", x = 2, y = 0.5, label = "(a) A. spurrelli", fontface= "italic")
gs
```

Just for fun, we could try putting on same plot somehow? maybe something like this? But Illustrator is best for these type of things!
```{r}
grid.newpage()
grid.draw(rbind(ggplotGrob(gs), ggplotGrob(gc), size = "last"))
```

















**********
Next let's load the data for the predation experiments!
```{r}
f <- curl("https://raw.githubusercontent.com/bguell/data-reanalysis-assignment/master/predation%20experiments.csv") #read in csv file from Github
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d) #see data and column names

#so this is a compilation of the snake and fish predation experiments! We'll have to subset data in order to only run models on each experiment and the predictors we chosse.

#Lets start with the snake predation experiment and data:
library(lme4) #require package to fun GLMM
library(car) #require package to allow for Anova on GLMM and effects of predictors
library(dplyr)
```

Interactive effect of age and species
```{r}
#lets filter out only observations for snake predation experiments
dt = filter(d, predation_type =="snake")
dt

m=glm(survival_prop~age*species, family = "binomial", weights = initial, data = dt)
summary(m)
Anova(m)
```

Effect of age on callidryas escape success:
```{r}

dc = filter(d, predation_type =="snake")
dc = filter(dc, species=="callidryas")

m=glm(survival_prop~age, family = "binomial", weights = initial, data = dc)
summary(m)
Anova(m)

```

Effect of age on spurrelli escape success:
```{r}
ds = filter(d, predation_type =="snake")
ds = filter(d, species=="spurrelli")

m=glm(survival_prop~age, family = "binomial", weights = initial, data = ds)
summary(m)
Anova(m)
```

